---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Introduction


Behavioral Risk Factor Surveillance System (BRFSS)

http://www.cdc.gov/brfss/

Percentage of US Adults who reported Fair or poor health
https://www.cdc.gov/chronicdisease/resources/publications/aag/brfss.htm
Can I generate a map?


Minnesota Initiatives
Minnesota collapsed
Provide data on current tobacco use to inform initiatives to prevent and reduce tobacco use in Minnesota.
Plan and evaluate Minnesota public health goals.
Assess exposure to environmental health risks in the home from lead, asbestos, radon, and well-water contamination.
Monitor the use of preventive health services and the prevalence of chronic conditions to evaluate the effectiveness of intervention strategies.

## Setup

```{r set-environment}
setwd('C:/Development/R Projects/Duke/IntroProbData/Course Project')
```

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
brfss <- load("brfss2013.RData")
```



* * *

## Part 1: Data


* * *

## Part 2: Research questions

**Research quesion 1:**

**Research quesion 2:**

**Research quesion 3:**


* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**

```{r}

```



**Research quesion 2:**

```{r}

```



**Research quesion 3:**

```{r}

```

