---
title: "Exploring the BRFSS data"
author: Scott Arendt
date: November 8 2016
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Behavioral Risk Factor Surveillance System (BRFSS)

The BRFSS is a collection of health related telephone surveys conducted annually.  There are 400,000 surveys performed each year on a wide variety of health issues and risk behaviors on adult U.S. residents.  The data is collected from all 50 states along with the District of Columbia.

The data can be found online at CDC website.  I am using the 2013 set of data for this analysis.

http://www.cdc.gov/brfss/

Percentage of US Adults who reported Fair or poor health
https://www.cdc.gov/chronicdisease/resources/publications/aag/brfss.htm
Can I generate a map?


Minnesota Initiatives

Provide data on current tobacco use to inform initiatives to prevent and reduce tobacco use in Minnesota.
Plan and evaluate Minnesota public health goals.
Assess exposure to environmental health risks in the home from lead, asbestos, radon, and well-water contamination.
Monitor the use of preventive health services and the prevalence of chronic conditions to evaluate the effectiveness of intervention strategies.

## Setup

```{r set-environment, echo=FALSE}
setwd('C:/Development/R Projects/Duke/IntroProbData/Course Project')
```

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data, echo=FALSE}
load("brfss2013.RData")
```



* * *

## Part 1: Data


* * *

## Part 2: Research questions

**Does the amount of sleep have any correlation to the general feeling of wellness?  Is there a difference between men and women?**

**Smoking and Income Correlation**

**Research quesion 3:**


* * *

## Part 3: Exploratory data analysis

**Research quesion 1:**

```{r}
# Build data representing the sleep time and general health.  Do I want to include physical and mental health days
# brfss2013 %>%
#   filter(!is.na(genhlth)) %>%
#   group_by(genhlth) %>%
#   summarise(count = n()) 
# 
# brfss2013 %>%
#   filter(!is.na(sleptim1)) %>%
#   filter(sleptim1<=24) %>%
#   group_by(as.numeric(sleptim1)) %>%
#   summarise(count = n()) %>%
#   str()

sleepANDhealth <-
  brfss2013 %>%
    filter(!is.na(genhlth)) %>%
    filter(!is.na(sleptim1)) %>%
    filter(sleptim1 <= 24) %>%
    select(genhlth, sleptim1)
```




**Research quesion 2:**

```{r}

```



**Research quesion 3:**

```{r}

```

## Appendix A - Code Book

I am listing a subset of the code book, describing only the columns that I used in my analysis.  It would be impractical to include the entire code book in this Appendix.

General Health
Section: 1.1 Health Status Type: Num
Column: 80 SAS Variable Name: GENHLTH
Prologue:
Description: Would you say that in general your health is:

Value Value Label
1     Excellent
2     Very good
3     Good
4     Fair
5     Poor
7     Don't know/Not Sure
9     Refused
BLANK Not asked or Missing


How Much Time Do You Sleep
Section: 4.1 Inadequate Sleep Type: Num
Column: 91-92 SAS Variable Name: SLEPTIM1
Prologue: I would like to ask you about your sleep pattern.
Description: On average, how many hours of sleep do you get in a 24-hour period?
Weighted
Value   Value Label
1 - 24  Number of hours [1-24]
77      Don't know/Not Sure
99      Refused 


Respondents Sex
Section: 8.21 Demographics Type: Num
Column: 178 SAS Variable Name: SEX
Prologue:
Description: Indicate sex of respondent.
Value 
Percentage
1   Male
2   Female



The complete code book is found on the CDC website here:
http://www.cdc.gov/brfss/annual_data/2013/pdf/codebook13_llcp.pdf